Emittance preservation of an electron beam in a loaded quasi-linear plasma wakefield

### 2017-10-05 Tex/Paper.tex
Some source cleanup and fixes. Added citations.

\documentclass[aps,prstab,reprint,amsmath,amssymb,groupedaddress,onecolumn]{revtex4-1}
\bibliographystyle{apsrev4-1}

@@

\begin{abstract}
We investigate beam loading and emittance preservation for a high-charge electron beam being accelerated in quasi-linear
plasma wakefield driven by a short proton beam. The structure of the studied wakefields are similar to those of a long, modulated
proton beam, such as the AWAKE proton driver.  We show that by %
properly choosing the electron beam parameters and exploiting two well known effects, beam loading of the wake field and full blow out
of plasma electrons by the accelerated beam, the electron beam can %PM be accelerated - EA ok
gain large amounts of energy with a narrow final energy spread (\%-level) and without significant emittance growth.
\end{abstract}

@@

Beam driven plasma wakefield accelerators \cite{chen:1985} have the potential to offer compact linear accelerators with high energy
gradients, and have been of interest for several decades. With a relativistic charged particle drive
beam travelling through a plasma, a strong wakefield is excited, that can be loaded by a trailing witness beam. When the
witness beam optimally loads the wakefield, an increase in absolute energy spread can be kept small. %to a minimum. %PM - EA ok
The concept
has been demonstrated experimentally in the past using electron drive beams accelerating electron witness beams
\cite{rosenzweig:1988, blumenfeld:2007, kallos:2008} [EA: M. Litos et al., Nature 515, 92?95 (2014)].

@@

AWAKE at CERN is a proof of concept proton beam driven plasma wakefield
accelerator experiment \cite{awake_collaboration:2014}, currently in its first phase of operation. The experiment uses a $400\unit{GeV}$ proton beam delivered by the
SPS as its driver, and a single $10\unit{m}$ plasma stage with a nominal plasma density of $7\nexp{14}\unit{cm}^{-3}$
\cite{gschwendtner:2016}. This plasma density corresponds to $\lambda_{pe} = 1.26\unit{mm}$ -- matched to the transverse
size of the SPS proton beam such that $k_{pe}\sigma_{x,y} = 1$, where $k_{pe} = 2\pi/\lambda_{pe}$ is the plasma wave
number, %PM. - EA ok
$\lambda_{pe}=2\pi c/\omega_{pe}$ and $\omega_{pe}=\left(n_0e^2/m_e\epsilon_0\right)^{1/2}$ is the plasma electron angular frequency. %

@@

The plans for AWAKE Run 2 proposes to use two plasma sections, illustrated in figure \ref{Fig:AWAKER2}. The first
section of about $4\unit{m}$ is the self-modulation stage where the proton beam undergoes self-modulation without the electron
beam present. The electron witness beam is then injected into the modulated proton beam before section two, where it
undergoes acceleration. The self-modulated proton beam does not produce a fully non-linear wakefield, and
 therefore not all plasma electrons are evacuated from the plasma bubble. The result is that the focusing force does not
increase linearly with radius %PM
and the accelerated beam emittance is not preserved.

@@

The focus in this study is the beam loading of the wakefields driven by the proton beam.   Studies of self-modulated proton bunches show that the proton bunch evolve as it propagates through a uniform plasma [EA: K. V. Lotov, Phys. Plas. 18, 024501 (2011], but that small variations in the plasma profile the modulation, and thus the wake fields, may be stabilized over long distances [EA: K. V. Lotov, Phys. Plas. 18, 024501 (2011), K. V. Lotov, Phys. Plas. 22, 103110 (2015), A. Caldwell and K. Lotov, Phys. Plas. 5, 103101 (2011)].  To study the witness beam evolution in a stable wake, independent of the dynamics of the self-modulation, %PMwe define here - EA ok
here we use a single, non-evolving proton bunch as driver.  The proton bunch parameters are chosen so that key features of the wake - the plasma electron density in the wake and the longitudinal electric field - are the same as %PM - EA ok
in the wake of a self-modulated proton beam with AWAKE baseline parameters \cite{gschwendtner:2016}.  Both the proton beam and the witness have Gaussian %PM - EA ok
longitudinal charge distribution and bi-Gaussian transverse charge distributions.

@@

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/plasmaDenTWake}
    \caption{\label{Fig:PlasmaDenTWake} %PMThe p
    Plasma density in grey with the drive beam (blue) and the witness beam (red)
        superimposed. The line plot indicates the transverse wakefield gradient $dW_{x}/dx$ where
        $W_{x} = E_{x} - v_{b} B_{y}$, evaluated along the beam axis.
        %PM
        Beams move to the left.}
\end{figure}

@@

%PM The baseline AWAKE drive beam current is insufficient to reach the non-linear regime and produce a plasma bubble. - EA ok
The modulation process used in AWAKE does not reach the fully non-linear regime and thus does not produce a bubble void of plasma electrons. When
the SPS beam, containing $3\nexp{11}$ protons \cite{gschwendtner:2016}, enters the second plasma stage (figure \ref{Fig:AWAKER2}), the peak
electric field is expected to be about $500\funit{MV}{m}$. The plasma electron density are only depleted to around $65\%$ of
nominal plasma density at the point where we inject the electron beam \cite{awake_collaboration:2016}.  The plasma electron density depletion and the peak field is replicated using a single %PM proton 
bunch with $1.46\nexp{10}$ protons (or $2.34\unit{nC}$), a %PM bunch
length $\sigma_{z} = 40\unit{\mu m}$ ($7\unit{kA}$), and a transverse size $\sigma_{x,y,proton} = 200\unit{\mu m}$.   The peak density is $0.83\cdot n_{0}$ results in a quasi-linear wake. %PM The beam energy is $400\unit{GeV}$.  
To avoid transverse evolution of the proton driver, emulating the stable propagation of the self-modulated proton beam [EA: K. V. Lotov, Phys. Plas. 18, 024501 (2011), K. V. Lotov, Phys. Plas. 22, 103110 (2015), A. Caldwell and K. Lotov, Phys. Plas. 5, 103101 (2011)] we freeze the transverse evolution of the equivalent proton bunch by increasing the particles mass by six orders of magnitude.

@@

%The small size of the accelerating structure produced by the self-modulated beam also
%puts constraints on the size of the electron witness beam, and thus its charge and current, if at the same time we wish
%to prevent large energy spread.  
%PM not sure what the point of the above sentence is ... - EA ok
In order to prevent large amplitude oscillations %PM
of the witness beam particles, which may cause additional energy spread as well as emittance growth, we %PMstudy
consider a witness beam matched to the plasma density.  
%The beam matching relation is given by
%\begin{equation}
%    \beta = \frac{\sigma_{x,y}^{2}\gamma}{\epsilon_{N}}
%          = \frac{c}{\omega_{pe}}\sqrt{2\gamma}, \label{EQ:MatchedB}
%    % = c\sqrt{\frac{m_{e}\varepsilon_{0}}{n_{pe}e^{2}}2\gamma}
%\end{equation}
%where $\beta$ is the Twiss parameter. It then follows that the transverse profile of the beam must satisfy
%PM Not sure what the point of the above lines is ....
The matched beam transverse size is:
\begin{equation}
%    \sigma_{x,y}^{2} = c\epsilon_{N}\sqrt{2\frac{m_{e}\varepsilon_{0}}{n_{pe}e^{2}\gamma}} \label{EQ:MatchedR}
%    \sigma_{x,y} =\left(c^2\epsilon_{N}^2m_{e}\varepsilon_{0}n_{pe}e^{2}\gamma \right)^{1/4} \label{EQ:MatchedR} - % PM - EA: not correct
	\sigma_{x,y}=\left(\frac{2c^{2}\varepsilon_{\mathrm{\mathrm{N}}}^{2}m_{e}\varepsilon_{0}}{n_{pe}e^{2}\gamma}\right)^{1/4}  \label{EQ:MatchedR}  % EA 
\end{equation}

We %furthermore PM - EA ok
assume an initial normalized emittance of $\epsilon_{N} = 2\unit{\mu m}$. This emittance is possible to produce with a standard RF-injector, while at the same time yielding a sufficient narrow beam. % EA: comment, perhaps put a ref here [***ref].

Beam loading of a short witness beam is sensitive to its position relative to the electric field \cite{tzoufras:2009} %PM
as well as to its dephasing with respect to the wakefields when at low energy.
%PM A low gamma witness beam can easily drift out of phase with the drive beam wakefield. The drift of the beams is proportional to $\gamma^{-2}$,  
To eliminate %initial 
de-phasing of the witness beam, % PM
the initial beam energy %PMwas
is set such
that $\gamma_{eb} = \gamma_{pb} = 426.3$, giving an energy of $217\unit{MeV}$.  A lower initial energy is likely to be
sufficient for AWAKE Run 2 injection. 
% EA: we have studied this to some extent, but unfortunately we do not have a good refence for the last statement, no details in \cite{berglyd_olsen:2015}.

Eq. (\ref{EQ:MatchedR}) yields a transverse size of $\sigma_{x,y}$ of $5.25\unit{\mu m}$, narrow compared to the drive beam $\sigma_{x,y,proton} = 200\unit{\mu m}$.   The bunch length was set to $\sigma_{z}=60\unit{\mu m}$, based on earlier %PM
beam loading studies \cite{berglyd_olsen:2016}.  The charge %PM was
is adjusted to $100\unit{pC}$ for optimal beam loading, as discussed in the next section. We refer to the defined drive beam and witness beam parameter set as the \emph{base case}.  
% EA: the key figure, figure 2, was not discussed in the text until very far down!  I therefore added (text could possibly be improved)
Figure \ref{Fig:PlasmaDenTWake} shows the two beams from QuickPIC simulations of the initial time step, for the base case parameters; the proton bunch in blue, the trailing electron beam in red and the plasma electron density in grey.

@@

The relatively small size of the witness beam %PM compared to the proton beam put some restrictions on the transverse resolution of the witness beam as we need a simulation box wide enough to contain the larger beam while resolving the narrower one (see figure \ref{Fig:PlasmaDenTWake}). 
puts constraints on the transverse cell size and number in the simulations: small size to resolve the small electron beam and large number of to resolve the large proton beam and the wakefields. %
We use %PMd  - EA ok
a transverse grid cell size of $1.17\unit{\mu m}$, and of
$2.34\unit{\mu m}$ for the longitudinal grid cells for the simulations presented in section \ref{S:BL}. The witness beam
was simulated with $16.8\nexp{6}$ non-weighted particles.  ***NN*** plasma particles were simulated per grid cell.
%   EA: Veronica to fill in NN above

@@

\begin{figure*}[hbt]
    \includegraphics[width=0.495\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamEmittance}
    \includegraphics[width=0.495\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamEmittanceOffset}
    \caption{\label{Fig:BeamEmitt} Beam density in blue along the beam axis for an on-axis beam with respect to the
        drive beam axis (left), and an offset beam (right) with an offset of one $\sigma_{x} = 5.24\unit{\mu m}$ in the
        x-plane. At four different positions $z$ in the plasma stage. The red lines show a moving window calculation of
        transverse normalised emittance. The moving window calculation uses longitudinal slices of
        $l = 4\times\Delta\xi = 9.38\unit{\mu m}$ with a step of $\Delta\xi$. Only slices with more than $100$ macro
        particles have been included. The plasma density profile is included in green, and scaled up by a factor of
        $100$ to be visible. These simulations were run with an LHC energy drive beam of $7\unit{TeV}$}
\end{figure*}

\begin{figure*}[hbt]
    \includegraphics[width=\linewidth,trim={0mm 0mm 0mm 0mm},clip]{figures/beamFilamentationAll}
    \caption{\label{Fig:BeamFilament} Plots \textbf{a} to \textbf{f} show the transverse phase space of the offset
        electron beam at different plasma positions. Plot \textbf{g} shows the macro particle mean position, and plot
        \textbf{h} their RMS spread. Plots \textbf{a}, \textbf{b} and \textbf{c}, as well as the blue lines in plots
        \textbf{g} and \textbf{h} represent particles %PM
        not in the ion column (see Fig.~\ref{Fig:BeamEmitt}), with position $1.40\unit{\mu m} < \xi < 1.42\unit{\mu m}$. Plots
        \textbf{d}, \textbf{e} and \textbf{f}, and the red lines in plots \textbf{g} and \textbf{h} represent particles %PM
        in the ion column 
        with position $1.55\unit{\mu m} < \xi < 1.57\unit{\mu m}$.}
\end{figure*}

@@

%~ Our configuration also shows some robustness to small electron beam transverse offsets as the bubble generated by the
%~ witness beam wakefield follows the beam itself.
%PM
So far we have considered a witness beam injected on the axis of the proton bunch. %
We now briefly examine the case of injection of a witness beam with an offset with respect to the proton bunch axis. %
Since the witness beam creates its own plasma bubble, the emittance of the part of the beam inside the
bubble is not affected by small transverse offsets of the witness beam with respect to the proton beam axis. This is illustrated in figure \ref{Fig:BeamEmitt}, right, for an electron beam offset of $1\sigma_{xe}$. Emittance preservation for small offsets is an added benefit of this 
accelerating regime, and may ease transverse injection tolerances. 
The head of the beam experiences a 
%larger initial emittance growth (see figure \ref{Fig:BeamFilament}a-f),
%PM
%THERE IS NO EMITTANCE SHOWN OR NOTHING TO COMPARE TO ... WHAT TO DO????
% EA: one way out, can compare emittance growth using fig 5 instead, as follows (alternatively, a new set of figures must be created)
larger initial emittance growth than for the on-axis case (compare figure \ref{Fig:BeamEmitt}, left, to \ref{Fig:BeamEmitt}, right), however, also for the head the emittance growth ends after the first few meters. Figure \ref{Fig:BeamFilament} a-c show the phase space of the head of the electron beam after 0, 1.0 and 2.5 m, while figure \ref{Fig:BeamFilament} d-f show the phase space of the trailing part of the beam.  The centroid oscillations of the head and the trailing part are shown in figure \ref{Fig:BeamFilament} g.  This effect of a transverse offset is likely to be greater for larger offsets as the beam oscillates
around the axis of the drive beam wakefield.% EA: would be good to include a statement of how large the offsets can be before when the electron beam starts to break up / not preserve emittance.

The transverse beam size within the bubble, where normalized emittance is preserved, follows the evolution given by
equation \ref{EQ:MatchedR} %PM where the plasma density is given by the ion density - EA ok
,i .e., evolves to stay matched. The on-axis density of the electron
beam, as a result, increases as its gamma factor increases and its transverse size decreases. The effect can be seen in
figure \ref{Fig:BeamFilament} h. This has the potential to cause overloading of the field. However, for the base case
no significant overloading %PMwas
is observed.

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQuality}
    \caption{\label{Fig:BeamQ} Ratio of total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a
        function of initial beam charge (blue %PM
        symbols, lines), and relative energy spread of the accepted charge (red %PM
        symbols, lines), after
        $4\unit{m}$ of plasma and with an initial emittance $\epsilon_{N,0}=2\unit{\mu m}$. These are shown for four
        different $\sigma_{z}$ from $40\unit{\mu m}$ to $100\unit{\mu m}$. The detailed studies presented in beam
        loading section correspond to the square marked lines at $100\unit{pC}$.}
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQualityAbs}
    \caption{\label{Fig:BeamQAbs} Total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a function of
        initial beam charge (blue %PM
        symbols, lines), and final %PMmomentum
        energy (red %PM
        symbols, lines), after $4\unit{m}$ of plasma and with an initial emittance
        $\epsilon_{N,0}=2\unit{\mu m}$.}
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQualityEmittance}
    \caption{\label{Fig:BeamQEmit} Ratio of total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a
        function of initial emittance (blue %PM
        symbols, lines), and relative energy spread of the accepted charge (red %PM
        symbols, lines), after $4\unit{m}$
        of plasma.}
\end{figure}

@@

Figure \ref{Fig:BeamQ}  show the dependence of %PM emittance
charge and energy spread on bunch length and %PMbunch
incoming charge.  An initial beam
emittance of $2\unit{\mu m}$ was used.  %
%PM
Therefore, we define the fractional charge as that whose emittance remains smaller than $2.1\unit{\mu m}$. %PM We ran the scan with beam charges from 
The beam charge range $10\unit{pC}$ to $300\unit{pC}$ and %with length
the $\sigma_{z}$ range from $40\unit{\mu m}$ to $100\unit{\mu m}$.  As can be seen from figure \ref{Fig:BeamQ} %, red curves
(red curves), both the
$40\unit{\mu m}$ and the $60\unit{\mu m}$ beam %PMhas
have a well defined minimum energy spread with a beam charge
$\approx 50\unit{pC}$ and $\approx 100\unit{pC}$, respectively. Lower beam charges tend to under-load the electric field,
while higher beam charges tend to overload it. It is also clear that longer beams with respect to the accelerating phase
of the field, $\approx\lambda_{pe}/4$, %PMwill
does not optimally load the wake, thus producing a larger spread in energy.  The blue curves show the fraction of %PMthe sliced emittance where the
charge whose initial emittance is preserved (the slice%Md
  emittance for the base case is shown in figure \ref{Fig:BeamEmitt}).  As the witness beam charge increases the fraction of slices with preserved emittance increases, as expected from an earlier onset of the bubble formation and also increase in the bubble size \cite{lu:2006-1, lu:2006}.
%PM SHOULD WE ADD SOMETHING LIKE:
We note here that operation with $100\unit{pC}$ leads to a significantly larger charge (factor $\sim 4$) with emittance preserved, at the expense of an increase of relative energy spread by a factor of two. %???

@@

Parameter studies indicate that up to a few $100\unit{pC}$, up to 70\% of the incoming bunch charge, is accelerated, for bunch of lengths of $40-60\unit{\mu m}$. Such electron bunches may be generated by an injector based on a standard RF photo-emission gun [EA: S. Doebert, private communication].

>>>>

\documentclass[aps,prstab,reprint,amsmath,amssymb,groupedaddress]{revtex4-1}
\bibliographystyle{apsrev4-1}

@@

\begin{abstract}
We investigate beam loading and emittance preservation for a high-charge electron beam being accelerated in quasi-linear
plasma wakefield driven by a short proton beam. The structure of the studied wakefields are similar to those of a long,
modulated proton beam, such as the AWAKE proton driver. We show that by properly choosing the electron beam parameters
and exploiting two well known effects, beam loading of the wake field and full blow out of plasma electrons by the
accelerated beam, the electron beam can %PM be accelerated - EA ok
gain large amounts of energy with a narrow final energy spread (\%-level) and without significant emittance growth.
\end{abstract}

@@

Beam driven plasma wakefield accelerators \cite{chen:1985} have the potential to offer compact linear accelerators with
high energy gradients, and have been of interest for several decades. With a relativistic charged particle drive
beam travelling through a plasma, a strong wakefield is excited, that can be loaded by a trailing witness beam. When the
witness beam optimally loads the wakefield, an increase in absolute energy spread can be kept small. %to a minimum. %PM - EA ok
The concept has been demonstrated experimentally in the past using electron drive beams accelerating electron witness
beams \cite{rosenzweig:1988, blumenfeld:2007, kallos:2008, litos:2014}.

@@

AWAKE at CERN is a proof of concept proton beam driven plasma wakefield accelerator experiment
\cite{awake_collaboration:2014}, currently in its first phase of operation. The experiment uses a $400\unit{GeV}$ proton
beam delivered by the SPS as its driver, and a single $10\unit{m}$ plasma stage with a nominal plasma density of
$7\nexp{14}\unit{cm}^{-3}$ \cite{gschwendtner:2016}. This plasma density corresponds to $\lambda_{pe} = 1.26\unit{mm}$
-- matched to the transverse size of the SPS proton beam such that $k_{pe}\sigma_{x,y,pb} = 1$, where
$k_{pe} = 2\pi/\lambda_{pe}$ is the plasma wave number, %PM. - EA ok
$\lambda_{pe}=2\pi c/\omega_{pe}$ and $\omega_{pe}=\left(n_0e^2/m_e\epsilon_0\right)^{1/2}$ is the plasma electron
angular frequency.

@@

The plans for AWAKE Run 2 proposes to use two plasma sections, illustrated in figure \ref{Fig:AWAKER2}. The first
section of about $4\unit{m}$ is the self-modulation stage where the proton beam undergoes self-modulation without the
electron beam present. The electron witness beam is then injected into the modulated proton beam before section two,
where it undergoes acceleration. The self-modulated proton beam does not produce a fully non-linear wakefield, and
therefore not all plasma electrons are evacuated from the plasma bubble. The result is that the focusing force does not
increase linearly with radius and the accelerated beam emittance is not preserved.

@@

The focus in this study is the beam loading of the wakefields driven by the proton beam. Studies of self-modulated
proton bunches show that the proton bunch evolve as it propagates through a uniform plasma \cite{lotov:2011}, but that
small variations in the plasma profile the modulation, and thus the wake fields, may be stabilized over long distances
\cite{lotov:2011, lotov:2015, caldwell:2011}.  To study the witness beam evolution in a stable wake, independent of the
dynamics of the self-modulation, %PMwe define here - EA ok
here we use a single, non-evolving proton bunch as driver. The proton bunch parameters are chosen so that key features of
the wake -- the plasma electron density in the wake and the longitudinal electric field -- are the same as %PM - EA ok
in the wake of a self-modulated proton beam with AWAKE baseline parameters \cite{gschwendtner:2016}. Both the proton
beam and the witness have Gaussian %PM - EA ok
longitudinal charge distribution and bi-Gaussian transverse charge distributions.

@@

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/plasmaDenTWake}
    \caption{\label{Fig:PlasmaDenTWake} Plasma density in grey with the drive beam (blue) and the witness beam (red)
        superimposed. The line plot indicates the transverse wakefield gradient $dW_{x}/dx$ where
        $W_{x} = E_{x} - v_{b} B_{y}$, evaluated along the beam axis. Beams move to the left.}
\end{figure}

@@

%PM The baseline AWAKE drive beam current is insufficient to reach the non-linear regime and produce a plasma bubble. - EA ok
The modulation process used in AWAKE does not reach the fully non-linear regime and thus does not produce a bubble void
of plasma electrons. When the SPS beam, containing $3\nexp{11}$ protons \cite{gschwendtner:2016}, enters the second
plasma stage (figure \ref{Fig:AWAKER2}), the peak electric field is expected to be about $500\funit{MV}{m}$. The plasma
electron density are only depleted to around $65\%$ of nominal plasma density at the point where we inject the electron
beam \cite{awake_collaboration:2016}.  The plasma electron density depletion and the peak field is replicated using a
single %PM proton 
bunch with $1.46\nexp{10}$ protons (or $2.34\unit{nC}$), a %PM bunch
length $\sigma_{z} = 40\unit{\mu m}$ ($7\unit{kA}$), and a transverse size $\sigma_{x,y,pb} = 200\unit{\mu m}$.
The peak density is $0.83\cdot n_{0}$ results in a quasi-linear wake. %PM The beam energy is $400\unit{GeV}$.
To avoid transverse evolution of the proton driver, emulating the stable propagation of the self-modulated proton beam
\cite{lotov:2011, lotov:2015, caldwell:2011} we freeze the transverse evolution of the equivalent proton bunch by
increasing the particles mass by six orders of magnitude.

@@

%The small size of the accelerating structure produced by the self-modulated beam also
%puts constraints on the size of the electron witness beam, and thus its charge and current, if at the same time we wish
%to prevent large energy spread.  
%PM not sure what the point of the above sentence is ... - EA ok
In order to prevent large amplitude oscillations of the witness beam particles, which may cause additional energy spread
as well as emittance growth, we %PMstudy
consider a witness beam matched to the plasma density.  
%The beam matching relation is given by
%\begin{equation}
%    \beta = \frac{\sigma_{x,y}^{2}\gamma}{\epsilon_{N}}
%          = \frac{c}{\omega_{pe}}\sqrt{2\gamma}, \label{EQ:MatchedB}
%    % = c\sqrt{\frac{m_{e}\varepsilon_{0}}{n_{pe}e^{2}}2\gamma}
%\end{equation}
%where $\beta$ is the Twiss parameter. It then follows that the transverse profile of the beam must satisfy
%PM Not sure what the point of the above lines is ....
The matched beam transverse size is:
\begin{equation}
    \sigma_{x,y,eb}=\left(\frac{2c^{2}\epsilon_{\mathrm{N}}^{2}m_{e}\varepsilon_{0}}{n_{pe}e^{2}\gamma}\right)^{1/4}
    \label{EQ:Matched}
\end{equation}

We %furthermore PM - EA ok
assume an initial normalized emittance of $\epsilon_{\mathrm{N}} = 2\unit{\mu m}$. This emittance is possible to produce
with a standard RF-injector, while at the same time yielding a sufficient narrow beam. % EA: comment, perhaps put a ref here [***ref].

Beam loading of a short witness beam is sensitive to its position relative to the electric field \cite{tzoufras:2009}
as well as to its dephasing with respect to the wakefields when at low energy.
%PM A low gamma witness beam can easily drift out of phase with the drive beam wakefield. The drift of the beams is proportional to $\gamma^{-2}$,  
To eliminate %initial 
de-phasing of the witness beam, % PM
the initial beam energy %PMwas
is set such
that $\gamma_{eb} = \gamma_{pb} = 426.3$, giving an energy of $217\unit{MeV}$.  A lower initial energy is likely to be
sufficient for AWAKE Run 2 injection. 
% EA: we have studied this to some extent, but unfortunately we do not have a good refence for the last statement, no details in \cite{berglyd_olsen:2015}.

Eq. (\ref{EQ:Matched}) yields a transverse size of $\sigma_{x,y,eb}$ of $5.25\unit{\mu m}$, narrow compared to the drive
beam $\sigma_{x,y,pb} = 200\unit{\mu m}$. The bunch length was set to $\sigma_{z}=60\unit{\mu m}$, based on earlier
beam loading studies \cite{berglyd_olsen:2016}. The charge %PM was
is adjusted to $100\unit{pC}$ for optimal beam loading, as discussed in the next section. We refer to the defined drive
beam and witness beam parameter set as the \emph{base case}.  
% EA: the key figure, figure 2, was not discussed in the text until very far down!  I therefore added (text could possibly be improved)
Figure \ref{Fig:PlasmaDenTWake} shows the two beams from QuickPIC simulations of the initial time step, for the base
case parameters; the proton bunch in blue, the trailing electron beam in red and the plasma electron density in grey.

@@

The relatively small size of the witness beam %PM compared to the proton beam put some restrictions on the transverse resolution of the witness beam as we need a simulation box wide enough to contain the larger beam while resolving the narrower one (see figure \ref{Fig:PlasmaDenTWake}). 
puts constraints on the transverse cell size and number in the simulations: small size to resolve the small electron
beam and large number of to resolve the large proton beam and the wakefields.
We use %PMd  - EA ok
a transverse grid cell size of $1.17\unit{\mu m}$, and of $2.34\unit{\mu m}$ for the longitudinal grid cells for the
simulations presented in section \ref{S:BL}. The witness beam was simulated with $16.8\nexp{6}$ non-weighted particles.
%  ***NN*** plasma particles were simulated per grid cell.
% EA: Veronica to fill in NN above
% VO: QuickPIC doesn't specify particles per cell. Only in Osiris,

@@

\begin{figure*}[hbt]
    \includegraphics[width=0.495\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamEmittance}
    \includegraphics[width=0.495\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamEmittanceOffset}
    \caption{\label{Fig:BeamEmitt} Beam density in blue along the beam axis for an on-axis beam with respect to the
        drive beam axis (left), and an offset beam (right) with an offset of one $\sigma_{x,eb} = 5.24\unit{\mu m}$ in
        the x-plane. At four different positions $z$ in the plasma stage. The red lines show a moving window calculation
        of transverse normalised emittance. The moving window calculation uses longitudinal slices of
        $l = 4\times\Delta\xi = 9.38\unit{\mu m}$ with a step of $\Delta\xi$. Only slices with more than $100$ macro
        particles have been included. The plasma density profile is included in green, and scaled up by a factor of
        $100$ to be visible. These simulations were run with an LHC energy drive beam of $7\unit{TeV}$}
\end{figure*}

\begin{figure*}[hbt]
    \includegraphics[width=\linewidth,trim={0mm 0mm 0mm 0mm},clip]{figures/beamFilamentationAll}
    \caption{\label{Fig:BeamFilament} Plots \textbf{a} to \textbf{f} show the transverse phase space of the offset
        electron beam at different plasma positions. Plot \textbf{g} shows the macro particle mean position, and plot
        \textbf{h} their RMS spread. Plots \textbf{a}, \textbf{b} and \textbf{c}, as well as the blue lines in plots
        \textbf{g} and \textbf{h} represent particles %PM
        not in the ion column (see Fig.~\ref{Fig:BeamEmitt}), with position $1.40\unit{\mu m} < \xi < 1.42\unit{\mu m}$.
        Plots \textbf{d}, \textbf{e} and \textbf{f}, and the red lines in plots \textbf{g} and \textbf{h} represent
        particles in the ion column with position $1.55\unit{\mu m} < \xi < 1.57\unit{\mu m}$.}
\end{figure*}

@@

%~ Our configuration also shows some robustness to small electron beam transverse offsets as the bubble generated by the
%~ witness beam wakefield follows the beam itself.
%PM
So far we have considered a witness beam injected on the axis of the proton bunch. We now briefly examine the case of
injection of a witness beam with an offset with respect to the proton bunch axis. Since the witness beam creates its own
plasma bubble, the emittance of the part of the beam inside the bubble is not affected by small transverse offsets of
the witness beam with respect to the proton beam axis. This is illustrated in figure \ref{Fig:BeamEmitt}, right, for an
electron beam offset of $1\sigma_{x,eb}$. Emittance preservation for small offsets is an added benefit of this
accelerating regime, and may ease transverse injection tolerances.  The head of the beam experiences a 
%larger initial emittance growth (see figure \ref{Fig:BeamFilament}a-f),
%PM
%THERE IS NO EMITTANCE SHOWN OR NOTHING TO COMPARE TO ... WHAT TO DO????
% EA: one way out, can compare emittance growth using fig 5 instead, as follows (alternatively, a new set of figures must be created)
larger initial emittance growth than for the on-axis case (compare figure \ref{Fig:BeamEmitt}, left, to
\ref{Fig:BeamEmitt}, right), however, also for the head the emittance growth ends after the first few meters.
Figure \ref{Fig:BeamFilament} a-c show the phase space of the head of the electron beam after $0$, $1.0$ and
$2.5\unit{m}$, while figure \ref{Fig:BeamFilament}d-f show the phase space of the trailing part of the beam. The
centroid oscillations of the head and the trailing part are shown in figure \ref{Fig:BeamFilament}g. This effect of a
transverse offset is likely to be greater for larger offsets as the beam oscillates around the axis of the drive beam
wakefield.
% EA: would be good to include a statement of how large the offsets can be before when the electron beam starts to break
%     up / not preserve emittance.

The transverse beam size within the bubble, where normalized emittance is preserved, follows the evolution given by
equation \ref{EQ:Matched} %PM where the plasma density is given by the ion density - EA ok
,i .e., evolves to stay matched. The on-axis density of the electron beam, as a result, increases as its gamma factor
increases and its transverse size decreases. The effect can be seen in figure \ref{Fig:BeamFilament}h. This has the
potential to cause overloading of the field. However, for the base case no significant overloading %PMwas
is observed.

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQuality}
    \caption{\label{Fig:BeamQ} Ratio of total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a
        function of initial beam charge (blue symbols, lines), and relative energy spread of the accepted charge (red
        symbols, lines), after $4\unit{m}$ of plasma and with an initial emittance $\epsilon_{N,0}=2\unit{\mu m}$. These
        are shown for four different $\sigma_{z}$ from $40\unit{\mu m}$ to $100\unit{\mu m}$. The detailed studies
        presented in beam loading section correspond to the square marked lines at $100\unit{pC}$.}
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQualityAbs}
    \caption{\label{Fig:BeamQAbs} Total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a function of
        initial beam charge (blue symbols, lines), and final %PMmomentum
        energy (red symbols, lines), after $4\unit{m}$ of plasma and with an initial emittance
        $\epsilon_{\mathrm{N},0}=2\unit{\mu m}$.}
\end{figure}

\begin{figure}[hbt]
    \includegraphics[width=\linewidth,trim={2mm 0mm 2mm 0mm},clip]{figures/beamQualityEmittance}
    \caption{\label{Fig:BeamQEmit} Ratio of total beam charge with an emittance growth $\Delta\epsilon \leq 5\%$ as a
        function of initial emittance (blue symbols, lines), and relative energy spread of the accepted charge (red
        symbols, lines), after $4\unit{m}$ of plasma.}
\end{figure}

@@

Figure \ref{Fig:BeamQ}  show the dependence of %PM emittance
charge and energy spread on bunch length and %PMbunch
incoming charge.  An initial beam emittance of $2\unit{\mu m}$ was used. Therefore, we define the fractional charge as
that whose emittance remains smaller than $2.1\unit{\mu m}$. %PM We ran the scan with beam charges from 
The beam charge range $10\unit{pC}$ to $300\unit{pC}$ and %with length
the $\sigma_{z}$ range from $40\unit{\mu m}$ to $100\unit{\mu m}$.  As can be seen from figure \ref{Fig:BeamQ} %, red curves
(red curves), both the $40\unit{\mu m}$ and the $60\unit{\mu m}$ beam %PMhas
have a well defined minimum energy spread with a beam charge $\approx 50\unit{pC}$ and $\approx 100\unit{pC}$,
respectively. Lower beam charges tend to under-load the electric field, while higher beam charges tend to overload it.
It is also clear that longer beams with respect to the accelerating phase of the field, $\approx\lambda_{pe}/4$, %PMwill
does not optimally load the wake, thus producing a larger spread in energy. The blue curves show the fraction of
%PMthe sliced emittance where the
charge whose initial emittance is preserved (the slice emittance for the base case is shown in figure
\ref{Fig:BeamEmitt}). As the witness beam charge increases the fraction of slices with preserved emittance increases, as
expected from an earlier onset of the bubble formation and also increase in the bubble size \cite{lu:2006-1, lu:2006}.
%PM SHOULD WE ADD SOMETHING LIKE:
We note here that operation with $100\unit{pC}$ leads to a significantly larger charge (factor $\sim 4$) with emittance
preserved, at the expense of an increase of relative energy spread by a factor of two. %???

@@

Parameter studies indicate that up to a few $100\unit{pC}$, up to $70\%$ of the incoming bunch charge, is accelerated,
for bunch of lengths of $40-60\unit{\mu m}$. Such electron bunches may be generated by an injector based on a standard
RF photo-emission gun \cite{doebert:corr}.
